---
title: "Aides Drug"
author: "kimberly Lemus Munoz"
date: "06-14-2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```



```{r eval=FALSE}
triUmp <- triUmp%>%
mutate(across(where(is.character), as.factor))
```


```{r}
females<-filter(triUmp,gender=="F")
```


```{r eval=FALSE}
females <- females%>%
mutate(across(where(is.character), as.factor))
```



# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

The data show us females who took the placebo for aids got less than the drug meant to prevent aids. Thier is less than 10% of the people tested got aids, but among those that did a higher percentage of them had taken the drug.




### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```

Based on the table a little over 58% of females that took the Aids drug still contracted AIDS, and 41% of the males took the placebo. Based on the results the females that had taken the preventative drug still interacted with the virus, leading to believe the drug did not work.




## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
We can see that the actually drug is doing more harm than taking the palbaoe.The p-value was low in which was 
getting these result make no different which 7 out of 10000
which didn't make a different.hypothesis p-value .0001
0.00006596



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Based on the table a little over 58% of females that took the Aids drug still contracted AIDS, and 41% of the males took the placebo. Based on the results the females that had taken the preventative drug still interacted with the virus, leading to believe the drug did not work.Which we can see the the p-value .0001 
0.000056 of the p-value.

