---
title: "Aides Drug"
author: "kimberly Lemus Munoz"
date: "06-14-2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```



```{r eval=FALSE}
triUmp <- triUmp%>%
mutate(across(where(is.character), as.factor))
```


```{r}
females<-filter(triUmp,gender=="F")
```


```{r eval=FALSE}
females <- females%>%
mutate(across(where(is.character), as.factor))
```



# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```



The plot shows that females who took the placebo for aids got it less than those that took the drug meant to prevent aids.There's less than 10% of the people tested got aids, but among those that did a higher percentage of them had taken the drug.




### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```


The table shows about 60% of the people  got the virus, took the drug; while about 40% of the people  got the virus only took the placebo. It tells us that the drug was not very effective the reason for it more people who took the drug got aids than those who took a placebo.






## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



This test shows that 210 people who took the drug contracted aids while there were 147 people who took the placebo and got aids. The chances of getting these type results  with no difference are 6.596 out of 10,000. This shows that there is a difference and also, that distinction is  the medication doesn't work since additional individuals that took the medication got help than those that took the placebo.




# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->



The bar chart shows that more females that took the drug contracted aids than the placebo with nearly 60% who got aids from taking the drug and about 40% taking the placebo. The numerical graph shows that out of 100 people who contracted aids 59 took the drug and 41 took the placebo. The fisher test told us that 210 people that took the drug got aids and 147 people that took the placebo got aids. The p test shows that there is a 7 out of 10,000 chance of getting these results if there was no difference. With all this data we can conclude that the drug does not prevent aids, in fact, it makes people more likely to contract aids.

